---
title: "Income Prediction Report"
author: "Avinash"
date: "29/11/2017"
output:
  html_document:
    keep_md: true
---

<style>

.middle {
text-align: center;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(readr)
library(rmarkdown)
library(yaml)
library(knitr)
```

## Question

Is it possible to correctly identify which `salary-class`(greater than or less than $50k) a person belongs to based on the features such as `age`, `workclass`, `final-weight`, `education`, `education-number`, `marital-status`, `occupation`, `relationship`, `race`, `sex`, `capital-gain`, `capital-loss`, `hours-per-week` and `native-country` ?

## Hypothesis

The hypothesis is that it should be able to accurately identify the `salary-class` of a person based on the given features.


## Raw Data Obtained

A quick look at the data obtained from the UCI Machine Learning Reposiotry:

```{r}
raw_data <- read.csv("../data/train.csv")
kable(head(raw_data, 2))
```

## Summary of Continuous Variables

Statistical summary of all the continious variables from the dataset:

```{r}
summary_df <- read.csv("../doc/data_summary.csv")
kable(summary_df)
```


## Data Visualizations

### Analysis of Age Variable

<div class = "middle">
<img src = "../doc/age.png" width = "400" height = "400">
<img src = "../doc/age_prop.png" width = "400" height = "400">
</div>

The number of people who earns more than $50,000 increases from an age of 25 to a maximum number of people at around the age of 35.
About 40% of the people aged around 50 earn above $50,000.


### Analysis of Hours Per Week Variable

<div class = "middle">
<img src = "../doc/hours.png" width = "400" height = "400">
<img src = "../doc/hours_prop.png" width = "400" height = "400">
</div>


About 40% of the people who work about 45-60 hours a week earn more than \$50,000. Only a small proportion of the people who work below 45 hours earn more than \$50,000. Strangely, about 20% of the people who work close to 0 hours a week earn more than \$50,000.

### Analysis of Sex Variable

<div class = "middle">
<img src = "../doc/sex.png" width = "400" height = "400">
<img src = "../doc/sex_prop.png" width = "400" height = "400">
</div>

The number of records for `Male` are double than that of `Female`. The proportion of female who earn more than $50,000 are less than one-third of that of the male population.

### Analysis of Education Variable

<div class = "middle">
<img src = "../doc/education.png" width = "400" height = "400">
<img src = "../doc/education_prop.png" width = "400" height = "400">
</div>

The ratio of people who earn more than \$50,000 increases with their level of education. The proportion of people who earn more than \$50,000 is the people who have `Doctorate` and it is closely followed by people who went to `Professional school`. Following this trend the proportions for people who has not completed High school is very low.

### Analysis of Marital Status Variable

<div class = "middle">
<img src = "../doc/marital_status.png" width = "400" height = "400">
<img src = "../doc/marital_status_prop.png" width = "400" height = "400">

</div>

About 15,000 people from the population have a marital status of `Married-civ-spouse` and about 45% of the people in this group earn more than \$50,000. The next largest group in this section is `Never-married`, closely followed by `Divorced`; and only a small proportion of these groups earn above \$50,000.

### Analysis of Occupation Variable

<div class = "middle">
<img src = "../doc/occupation.png" width = "400" height = "400">
<img src = "../doc/occupation_prop.png" width = "400" height = "400">
</div>

Among the larger groups, about 50% of the people woking as Professional Speciality and Executive Managerial earn above \$50,000. One of the larger groups with the lowest proportion is `Other-services`. Only a small proportion of less than 10% get paid more than \$50,000.

### Analysis of Family Relationship Variable

<div class = "middle">
<img src = "../doc/relationship.png" width = "400" height = "400">
<img src = "../doc/relationship_prop.png" width = "400" height = "400">
</div>

When the sample is classified based on their relation,a higher proportion of `Wives` earn more than $50,000 than any other relation. This is closely follow by `Husband`. However, it should be also noted that the number of wives in the sample are much smaller when compared to husbands.

### Analysis of Race Variable

<div class = "middle">
<img src = "../doc/race.png" width = "400" height = "400">
<img src = "../doc/race_prop.png" width = "400" height = "400">
</div>


The races with the higher proportion of salaries are `White` and `Asian-Pac-Islander` at about 25% and other races have a proportion of about 10%. But it should be noted that there are only a few hundred people in the `Asian-Pac-Islander` group.

### Analysis of Work Class Variable

<div class = "middle">
<img src = "../doc/work_class.png" width = "400" height = "400">
<img src = "../doc/work_class_prop.png" width = "400" height = "400">
</div>


Majority of the people work in `Private` sector and about one-fourth of the people earn above \$50,000. The highest proportion of people who earn above \$50,000 are Self Employed. This is closely followed by `Federal-gov` and then `State-gov`.

### Model Using Random Forests


```{r, echo=FALSE, message=FALSE, warning=FALSE}
cm <- read_csv("../doc/confusion_matrix.csv", col_names = FALSE)

colnames(cm) <- c("Actual Less than 50", "Actual More than 50")
rownames(cm) <- c("Predict Less than 50", "Predict More than 50")

kable(cm)

```

When a model is created using only the age and hours a person works per week then an accuracy of 88% is obtained but this is obtained by classifying everying person as earning less than $50,000.
